\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage{algorithm} %
\usepackage{algpseudocode}
\usepackage{amsfonts}

\usepackage{tcolorbox}
\newtcolorbox{greenbox}[1]{colback=green!5!white,colframe=green!75!black,fonttitle=\bfseries,title=#1}


\usepackage{amsthm}
\usepackage{bm}

\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

\newtheorem*{remark}{Remark}

\usepackage{amsmath}
\DeclareMathOperator*{\argmin}{argmin} % no space, limits underneath in displays



\setlength{\parindent}{0pt} % for getting rid of first line indentation


% ================================================
% =====================  BEGIN DOCUMENT 
% ================================================

\title{Financial Math}
\author{Juan Pablo Bello Gonzalez}
\date{January 2024}


\begin{document}
\section{Introduction to financial markets}

\textbf{Question} If you are the issuer (short position) of an option how do you set a a strike price on the contract for short/put?

\textbf{Question} Is the number of shares I can buy/sell for the strike price also stipulated on the contract?


\begin{definition}
\textbf{Risk-free interest rate} is the rate an investor can expect to earn on an investment that carries 0 risk. Why would lending money to the bank carry risk? because of the time value of money (TVM). Compensates investors for the time-value of money: Their money is tied up and so they can't put it to use and inflation rises.
\end{definition}

\begin{definition}
\textbf{Time Value of Money}
A sum of money is worth more now than the same sum will be worth in the future. Main reasons are \textbf{inflation}, and due to its earnings potential in the interim.
\end{definition}

\begin{definition}
In a company: \textit{Equity = Assets - Liabilities}. 
\begin{itemize}
\item \textbf{Assets}: Everything you own which has value: infrastructure, patents, bonds, stocs, commodities. They have a random value over time $S_t$. 
\item \textbf{Liabilities}: Future debt 
\item \textbf{Owner's Equity}: What the owners would get if they sold the company. 
\end{itemize}
\end{definition}

\begin{definition}
\textbf{Stock}: A share in ownership of a company's equity. 
\begin{itemize}
\item Dividend: payment to shareholders from profits. Some stocks don't have dividends. 
\end{itemize}
\end{definition}

\begin{definition}
\textbf{Security:} A tradable financial asset. A house, stock, etc. 
\end{definition}

\begin{definition}
\textbf{Derivative:} A "secondary" security. A contract that derives its value and risk from  particular security. On option is a derivative of a stock because it is a "bet" on the stock, it has "derived" value from the stock.
\end{definition}


\begin{definition}
\textbf{Option}: A financial derivative/contract stipulated today ($t=0$) that gives the buyer the right to buy or sell the underlying asset within a specified period up to maturity $t=T$ and at a specified strikeprice independent of the asset value $S_t$. Just like with any bet, you can bet on both all outcomes:
\begin{itemize}
\item Long position: held by the holder (=buyer) of an option
\item short position: held by the writer (=seller) of an option. There is usually an upfront capital you have to put down for the shorting of a stock, you don't just sign you'll give it back in the future and take the money out of thin air. 
\item In theory, the short and long position profits will always mirror each other since they are the only ones involved in the bet. There are hidden fees in practice.
\end{itemize}
\end{definition}

We use the following language when talking about an option: 

\begin{itemize}
\item Call: right to buy at strike price (profit from stock increase over strike-price)
\item Put: right to sell at strike price (profit from stock decrease under strike-price)
\item European type: At maturity T.
\item American type: Any-time between now and maturity pay.
\item Asian type: Average stock price evolution up-until maturity ***.
\item Strike-price K: specified at $t=0$. 
\end{itemize}


\subsubsection{European Call option}
I am confident the stock price will increase over one year from current price $S_0$ above the threshold $K$. I can enter the following \textbf{bet}: 
\begin{itemize}
\item At time $t=0$ (today) I pay the \textbf{price of the option} $C_0 \geq 0$ to enter the bet. 
\item At \textbf{maturity}, say $T=1$. I have the \textit{right but not the obligation} to buy the stock for the \textbf{strike price}. 
\item If at maturity $S_T > K$, I would choose to exercise this option, that way I can buy the option for less that what it is worth at the time ($K$) and then sell it for $S_t$. Payout is $S_T-K$, profit is $S_T- K - C_0$.
\item If at maturity $S_T < K$, then it doesn't make sense for me to exercise the option. So my payout is $0$ and profit is $-C_0$ i.e. $100\%$ of my principal.
\item $S_t$ is a random variable, we do not know the evolution of the stock price $S_t$ otherwise we'd be rich. 
\end{itemize}

The payout for the long position (owner of the option) at some future one period time is: 
\[ C_T = max\{S_t - K, 0\} \]

This is a random variable which is a function of $S_t$. 

Now the central questions is: 

\begin{center}
What is a fair price for $C_0$ i.e. the price of the contract?
\end{center}

We may also be interested in the evolution of the price of this derivative $C_t$ overtime. Maybe we want to sell it? or just calculate payouts. 

\begin{center}
What is a fair price for $C_t , t \in [0, T)$?
\end{center}

Under the Black-Scholes model there is a closed form $C_0$ and $C_t$. 

\subsection{Options}
All options carry their associated risk. 

Consider an \textbf{European Option}. We make money if we are right about our prediction ($S_T > K$ for the call, $S_T < K$ for the put) exactly at time $T$.

Predicting points is much harder and unreliable than predicting trends, especially if the stock price is relatively volatile (really big up and downs to relative to stock price). Say we have a call option and the stock price of BMW is increasing consistently for one year but then at maturity the owner of company decides to go hunt an endangered tiger which causes the $S_T$ to plummet bellow $K$.

To protect against such things we may look to bet on path-dependant options like the Asian or American Option. This way we bet on trends and not on points.  \textbf{We expect these to be less risky?}

\subsubsection{Shorting Stock}
When you "short" a stock, there is no derivatives involved. You borrow the \textit{shares} from the broker and sell them at time 0. You are counting on the stock price dropping to make a profit. If the stock has indeed dropped bellow the price you paid and you decide it's reached a minimum, you buy it at the current price and give back the shares you borrowed and sold from the broker. Theoretically, if the stock drops to zero your profit is 100\% of what you sold it for $S_0$ as you can pay nothing to buy it back and repay it to the broker. On the other hand, if the stock keeps rising above the strike price, your debt is unbounded.

\subsubsection{Put option}
Say you believe the stock is going to go down but you don't have the stomach to short it (cuz you can loose an unbounded amount of money) you can buy a put option. Then you have the right to sell high for the strikeprice if the price at maturity is lower. you buy from the market for $S_T$ and sell for $K$. You also don't have the obligation to sell, so you don't have to buy it. Your loss is bounded by 100\%  of $C_0$ if you choose not to exercise the option, which is most likely less than the stock price. 

Unlike with shorting, you don't borrow and sell the stock in advanced. 

\subsection{Example}

Let's say a stock starts at $50$ and then it can go to $80$ or $20$. We have respective put and call options which we can buy for $5$ at strike prices of 60\$ for the call and 40\$ for the put. 

\begin{center}
Profit
\begin{tabular}{||c c c c c||} 
 \hline
 Security & Upfront Capital & Call & Short & Put\\ [0.5ex] 
 \hline
 \hline
 80 & 30 & 15 & -30  & \$5 \\ 
 \hline
 20 & -30 & -5 & 30 & \$15 \\
 \hline
\end{tabular}
\end{center}

\begin{center}
\% gain and loss
\begin{tabular}{||c c c c||} 
 \hline
 Security & Upfront Capital & \% Gain & \% Loss \\ [0.5ex] 
 \hline
 \hline
 Stock & 50 & 60\% & 60\% \\ 
 \hline
 Call & 5 & 300\% & 100\% \\
 \hline
\end{tabular}
\end{center}

From this table we see that the call option gives us \textit{leverage}. Putting leverage on your bet = "multiplying your potential gains or losses"


\section{Probability Theory}
Random variable is a function from sample space to real line. 

A discrete RV is characterized by its probability \textbf{mass} function. A \textbf{pmf} $f_X$ is such that:
\[f_X : \Omega \rightarrow \mathbb{R_{\geq 0}} \]
\[\sum_{x \in \Omega} f_X(x)=1 \]


A continuous RV is characterized by its probability \textbf{distribution} function. A \textbf{pdf} $f_Y$ is such that:
\[f_Y : \Omega=\mathbb{R} \rightarrow \mathbb{R_{\geq 0}} \]
\[\int_{\Omega} f_Y(y) dy =1 \]
The sample space for cts rv's is often the reals. 

\subsection{Independence}
For a sequence of RV's $X_1, X_2, \dots X_n$there are two principal notions of independence: 
\begin{itemize}
\item Mutual independence: independence property satisfied for all subsets of ${X_i}_{1<i<n}$.** \textbf{This is the one we are interested in.}
\item Pairwise independence: independence property satisfied only for pairs.
\end{itemize}

\subsection{Normal distribution}
Lets say a stock price is $P_n$. Does it make sense to model/assume the unit increments are normal? ie. $P_n-P_{n-1} \sim N(0,1)$?  Well fixed increases represent a larger percentage increase for smaller values than for large values. To exemplify: if my stock goes from 1 to 2 it increased by 100\%, but if it goes from 400 to 401 then it barely increased. So this is saying that at smaller stock values there is more aggressive changes than for bigger ones. We don't want this for modelling stock prices. The changes should be relative to the current price. Maybe it's human nature that when we are buying a house in the 8 digits we care less about the latter digits, but when we are buying candy we care about single dollar increases because it means large relative increases. Also this assumption would make things simpler:
\[ P_n-P_{n-1} \sim N(0,1)  \rightarrow P_n \sim N(0, \sqrt{n})\]

Instead we want the relative unit increases to be normally distributed, ie:
\[ \frac{P_n-P_{n-1}}{P_n} \sim N(0,1)\]

And the million dollar question: If it is the case that the relative stock price unit increments follow a standard normal distribution, what is the distribution of the stock price $P_n$?

\subsection{Log-Normal distribution}



\section{Stochastic processes}
A \textit{stochastic process} is:
\begin{itemize}
\item \textbf{Defn(1):} A sequence of random variables indexed by time (e,j. $X_0, X_1, \dots$ for discrete time)
\item \textbf{Defn(2):} A probability distribution over a space of paths. It is easier to do some math with this definition. Here, the graph of a stock evolution is just one realization of the paths random variable. Note that in practice we may not have the joint distribution for real world time series data but we can construct things like random walks from known joint normals. 
\end{itemize}

The goal of studying stochastic processes is to observe the past up to time $t'$ and be able to make an intelligent prediction about the process in the future. Depending on the process this can be trivial or impossible and anywhere in between. 

If our stochastic process is: 
\[ X_t = t \text{, with probability 1} \]
It is trivial to predict the future from the past if my stochastic process is:

If our stochastic process is: 
\[ 
X_t = 
\begin{cases}
t \forall t \text{, with probability 1/2} \\
-t \forall t \text{, with probability 1/2} \\
\end{cases}
\] 
Even with randomness introuduced, we can still describe the exact evolution of $X_{t \geq t'}$ in this case if we know just one realization of $X_{t<t'}$. This is because the sp is dependent. 


If our stochastic process is: 
\[ 
X_t = 
\begin{cases}
t \text{, with probability 1/2} \\
-t \text{, with probability 1/2} \\
\end{cases}
\]
Here the past can tell us nothing about the future because the sp is random and independent.

In the study of sp's we are interested in three types of questions:
\begin{enumerate}
\item What are the dependencies in the sequence of value? (ie. will knowing the past will help me predict the future)
\item What is the long term behaviour? Does the sequence converge to something or just bounce around. Law of large numbers and CLT. 
\item What are the boundary events? Events of interest to me, like how often does a stock drop by 10pts. 
\end{enumerate}

\subsection{Discrete}
\subsubsection{Simple Random Walk}
Denote our "steps of the walk" as $Y_i$ \textbf{i.i.d} random variables with: 
\[ Y_i = 
\begin{cases}
1 \text{, with prob 1/2} \\
-1 \text{, with prob 1/2} \\
\end{cases}
\]
then the \textit{one-dimensional} \textbf{random walk} $X_t$ is: 
\[ X_0 = 0 \text{ and } X_t = \sum_{i=1}^t Y_i \forall t>0 \]
It is one dimensional analogous to a drunk guy walking frowards and backwards along a straight line. Diagonal Motzkin paths make for a clearer picture. 

How does the random walk behave as time grows?
What is the distribution at time $t$? This is CLT stuff, the variance will be $t$ and the mean will be zero. Why is the variance t? well that's the maximum height about zero that the walk can achieve at time $t$ and it is only achieved if all steps are in one direction.  

**** Consult CLM. 


\section{Model}
We \textit{assume}:

\begin{enumerate}
\item $(S_t)$ is a \textit{stochastic process} 
\begin{itemize}
\item \textbf{Defn(1):} A sequence of random variables indexed by time (e,j. $X_0, X_1, \dots$ for discrete time)
\item \textbf{Defn(2):} A probability distribution over a space of paths. It is easier to do some math with this definition. Here, the graph of a stock evolution is just one realization of the paths random variable. Note that in practice we may not have the joint distribution for real world time series data but we can construct things like random walks from known joint normals. 
\end{itemize}

\item $S_t$ is modelled discretely.
\begin{itemize}
\item Discrete time stochastic process: at discrete times $0 < t_1 < \dots t_N = 0$ we observe the respective stock prices $S_{t1} \dots S_{tN}$.
\item Continuous time sp: for continuous $0 < t < T$ we have $S_t$ is a function with solid lines \textbf{(does not have to be continuous itself though)}. Can't express continuity in computers because we can't express sets of  infinite cardinalities, like subsets of the real line, with finite resources. 
\end{itemize}
\item $V_t$ is the value of the option. It is a function of time $t$ and the value of the underlying stock at or up-to time $t$, so: $V_t = V(S_t,t )$.

\end{enumerate}


Denote by $V_t = V(S_t,t )$ the value of the option at time $t$. It depends on the value of the underlying asset $S_t$ and sometimes, like in the Asian option case, it depends on the entire evolution up to time $t$. For each time $t > 0$, $S_T$ is a random variable ... and I think $S_t$ and $S_j$ don't have the same distribution, that's what makes them exciting, we don't have the joint but we can infer things

Tading options is considered risky, we see how the payouts are all or nothing. trading options is considered  riskier than trading the stock. 

This si teh leverage effect: The relative loss in the option can be much more than the relative loss in teh asset. Becuae of the all or nothing payout in options, you loose 100\% of what you paid to get into the bet $C_0 = V_0$.

\[ |\frac{S_t - S_0}{S_0}| << |\frac{V_s-V_0}{V_0}|\]

On a call/put option, when do we get the ultimate payout as the buyer of the option? when the difference $S_t-C$ is largest 

Let's think about american options because the maximization of european options depends entireley on $S_T$. For american options, we get to decide when we want to cash out. 


\section{Risk free asset}
We assume our cash-accounts continuously compound at risk free rate $r \geq 0$.

Definition makes sense, in practice, how do we get the risk free rate? We read whatever the "US treasury" rate says. Well you need this r for your caclualtions, it doesn't matter much what it is since you want to get results independent of its realized constant value. 

Our first, most simple equation, 

\[ dB(\tau) = r  B(\tau) d\tau \]


\subsection{Put-Call parity}
If the equation does not hold there may be an arbitrage opportunity. 

\section{One period Binomial}

\subsubsection{Pricing Options}
We can value options by finding a replicating portfolio. 

Another approach is by computing the discounted expected payout under a special probability measure $\mathbb{Q}$ which we call \textbf{risk neutral} or \textbf{arbitrage free pricing measure}.

The idea with the latter is to find $\mathbb{Q}$ such that:
\[ \mathbb{E}^{\mathbb{Q}}(S_t) = e^{rT}S_0 \]

also 
\[S_0 = e^{-rT}\mathbb{E}^{\mathbb{Q}}\]

That is, today's price of the stock is equal to the discounted expected value of the stock price at time $T$ given by $S_T$.


Recall that the price of an option \textit{does not depend} on the real world measure $\mathbb{P}$. the probability measure $\mathbb{Q}$ is what the market prices imply the price should be so that no arbitrage is possible. It is not the true probability measure. 

\end{document}


