\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage{algorithm} %
\usepackage{algpseudocode}
\usepackage{amsfonts}

\usepackage{tcolorbox}
\newtcolorbox{greenbox}[1]{colback=green!5!white,colframe=green!75!black,fonttitle=\bfseries,title=#1}


\usepackage{amsthm}
\usepackage{bm}

\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

\newtheorem*{remark}{Remark}

\usepackage{amsmath}
\DeclareMathOperator*{\argmin}{argmin} % no space, limits underneath in displays



\setlength{\parindent}{0pt} % for getting rid of first line indentation


% ================================================
% =====================  BEGIN DOCUMENT 
% ================================================

\title{Forecasting}
\author{Juan Pablo Bello Gonzalez}
\date{January 2024}

\begin{document}



\maketitle

\section{Statistics stuff}
\begin{itemize}
\item Forecasting means prediction 
\item Predicting a single number tells you nothing because we are dealing with probabilities and randomness, so always produce a confidence interval.
\item Pick a model, then residual diagnostics, if they pass you can do inference (prediction, condifence intervals, p-values). 
\item The idea behind residual diagnostics is to see if the random component of your model is truly random and does not include deterministic patterns you should be using in the model.
\item $Cov(X_i, X_j) = 0$ for uncorrelated RVs
\item for two RVs, Independence is a stronger condition than non-correlation
\item WHAT ARE DEGREES OF FREEDOM?


\end{itemize}

\section{Introduction to Forecasting}

\begin{definition}
Time series: A sequence of random variables indexed by time $\{ X_t : t\in T \subset \mathbb{N}\}$. It is a special case of a stochastic process. 
\end{definition}

\begin{definition}
Stochastic process: a collection of random variables (they don't need to have the same distribution!) indexed by a set $I$. $I$ could be continuous or discrete.
\end{definition}

We often want to model time series data with \textit{non-stationary trends}:
\begin{itemize}
\item Seasonality
\item trending mean
\item Non-constant variance (fanning in our out pattern)
\end{itemize}

Ideally given a time series $\{X_1, X_2, ... X_n\}$ we would know the \textit{joint distribution} of such vectors. 
\[ F(x_1, x_2, \dots x_n) = P(X_1, X_2, \dots X_n)\]

This is because individual (marginal) distributions of $X_i$ don't tell you about the combined behaviour.

We want to study a stochastic process as a mathematical which is a sequence of RV's with some predictable distribution ie not just a vector of n random and unrelated numbers. As such, the existence of a stochastic process should always be questioned. Kolmogorov's theorem tells us about the existence of  discrete time stochastic process build from $X_i's$ as long as the $X_i$ satisfy the two xonsistency conditions. 

If all finite dimensional distributions exist, and they are consistent with each other under marginalization, then the stochastic process exists.  Then \textbf{we can use finite dimensional models ot define a general process}.*****

All finite dimensional distributions of $X_1, X_2, \dots X_n$ are those of all non-empty subsets of the sequence ie $\{ X_1, X_2, \dots X_n \}^*$ so it is impossible to have all of these in practice since, if we are sampling, we don't even have the distribution of a single $X_i$ and must do with estimates. 

Fortunately, in most application we can get by with knowing $E[X_t], E[X_t^2], E[X_tX_j]$. Which is equivalent to knowing $E[X_t], Var(X_t), Cov(X_t, X_j)$ for \textbf{ALL} indices.

\begin{definition}
\textbf{A time series model} for the observed data $\{x_t\}$ is a specification of the joint distributions or more often in practice, the means and covariances of a sequence of random variables $\{X_t\}$ of which $\{x_t\}$ is a realization. 
\end{definition}


\subsection{Zero-mean time series models}
\subsubsection{i.i.d Noise}
\begin{definition}
\textbf{iid Noise} a sequence $\{X_i : i = 1,2\dots n\}$ with: 
\begin{itemize}
\item  independence property is satisfied: joint = product of marginals.
\item  $E[X_i] = 0 \forall i$

\end{itemize} 
\end{definition}

As is the case with a discrete time stochastic process with independent components, the memory-less property is satisfied.  
\begin{proof}
\[ P(X_{n+h} \leq x | X_1 \leq x_1, \dots, X_1 \leq x_n ) = \frac{P(X_{n+h} \leq x)  F(x_1, \dots x_n)}{F(x_1, \dots x_n)}\]
\end{proof}
Note that in the nominator above, the probability of the intersection of events is equal to the individual probabilities multiplied by independence. 

Also note how we can have the individual $X_i$ be continuously distributed but discrete stochastic process (indexed by the naturals)

\subsubsection{White Noise }
Sequence of uncorrelated random variables $\{X_1, X_2, \dots \}$ with zero mean and finite variance. Denoted $X_i \sim WN(0, \sigma)$
\begin{itemize}
\item $E(X_i)=0$ 
\item $Var$
\item $Cov(X_i, X_j) = 0$ for uncorrelated 
\end{itemize}


\subsubsection{Random Walk}
\begin{definition}
A R.V which is a \textbf{series} of \textbf{iid noise OR White noise} distributed R.V.s $S_t = \sum_{i=1}^t X_i$. Usually starting at zero. 
\end{definition}

What is the relationship between \textit{white noise} and \textit{iid noise}?
Independence is a stronger condition than non-correlation. For independence we need to know joint distributions but for non-correlation we don't. So we'd rather work with white noise. 

\section{Classical decomposition}
In regression we are always modelling $Y|X$ and not just $Y$ since we are actually using the explanatory variable to fit $Y$.

The general framework is to remove the \textbf{trend}, \textbf{seasonality}, and \textbf{non-constant variance} patterns in the data  so that \textit{hopefully} we are left with a stationary process. Then model the detrendized and deseasonalized data. 

\subsection{Additive classical decomposition}
Fitting models with trend and seasonality:
\[ X_t = m_t +S_t +Y_t \]

where,
\begin{itemize}
\item $X_t$ is value of the process at time $t$
\item $S_t$ models the periodic deterministic component of $X_t$ at time $t$
\item $m_t$ is the TREND deterministic component. Doesn't have to be linear.
\item $Y_t$ is the stochastic component, white noise 
\end{itemize}


Note that the expected value of the process depends on time: $E[X_t] = E[S_t] + E[m_t]$. Particularly if the expectation for the trend component is non-zero as the expectation of periodic functions is often constant.

We don't need Fourier series to model periodic patterns (in discrete stochastic processes), we can figure out the period $d$ and then use $d-1$ indicator variables whose coefficients give the period height at each point of time. Let's assume the period for some time series is 12 corresponding to 12 months, we use $d-1=11$ indicator variables $X_1, X_2, \dots, X_11$ to represent:
\begin{equation}
x_1 = 
\{
    \begin{cases}
        1, & \text{if month is February}\\
        0, & \text{else}
    \end{cases}
\end{equation}
The reason we don't need as many \textbf{indicator variables} as the period is that we can measure the first point (January) with just trend $X_1 = m_t+Y_t$ and then have the remaining $d-1$ coefficients $\beta_j$ tell us the deviation of point $X_{i+1}$ relative to $X_1$ it is the baseline. 


We use regression for $m_t$ and $S_t$, nothing new so far. The trend $m_t$ need not follow a linear regression model, we can model it to be any regular function ie. $m_t = \beta_0 + \exp{\beta_1t}$.

If we have monthly data, the period need not be $12$, it might be biweekly if we are measuring something that is correlated to paydays. 

\section{Loss Functions/Decision Theory}
Measure the quality of the fit. As you imagine, there are several ways to measure "quality".
We need a measure to compare the fitted predicted values to the truth. 

Denote the loss function as $L(Y, f(X))$. It measures the error in estimating $Y$ with the estimate $f(X)$

The loss function \textbf{is a random variable}. The expected value of the loss function with respect to the \textbf{joint probability function} is called the \textbf{risk}:
\[ E_{X,Y}[L(Y, f(X))] \]

A couple of things here:

The expectation is over the joint distribution in theory, but in practice, since we are tryint to use $X$ to predict $Y$ we \textit{assume $X$ is given} so the expecation will be ralative to this distribution ie. 

\[ E_{Y|X}[L(Y, f(X)) | X =x] = \int_{\mathbb{R}}  P(Y|X=x) dL(Y, f(X))\]  

Even these conditional probabilities are hard to wrap your mind around, we have to assume their distribution, often to uniform when we consider the \textbf{empirical} risk estimate:

\[ \hat E[L(y_i, f(x_i))] = \sum_{\Omega} L(y_i, f(x_i))\]  

\subsubsection{Quadratic Loss/Squared error loss}
$L_{SE}(Y, (f(X)) = (Y-f(X))^2$ where $f(X) = \hat Y :=$ the random variable $Y$ estimated as a function of $X$.

The risk of this loss function also has a special name: \textbf{MSE}
\[ MSE(f) = E_{X,Y}[L_{SE}(Y, f(X))] \]

\begin{theorem}
\textbf{Conditional Expectation Theorem}: If $X,Y$ are two random variables with $E(Y)=\mu, Var(Y) < \infty$, then the function $f$ which minimizes $MSE(f)$ for $X,Y$ is given by the conditional expectation:
\[f(X) = E_{Y|X}(Y|X) \]
\end{theorem}

\begin{proof}
************************************
\begin{itemize}
\item let $f(X) = c$ and show that the constant $c$ that minimizes the risk is the mean of $Y$.
\item Assume $c(x)$ minimizes 
\end{itemize}
\end{proof}

Notes:
\begin{itemize}
\item $E(Y|X=x)$ is a number depending on x. So it's a function of x
\item $E(Y|X)$ is a random variable
\end{itemize}

\subsubsection{Absolute Error Loss}
Minimized by the median



\section{Regression}
The MSE is minimized by the conditional expection $E(Y|X)$. But we don't really know how to get this, so we take a step back and assume the form of $f(X)= \bm X \bm \beta + \epsilon$ in the risk minimization problem (called OLS, RSSQ) and then solve for the coefficients using matrix methods. 

In multiple regression we have some simplifying assumptions s.t. 
\[ f(X;\bm \beta) = E[Y|X] = \beta_0 + \sum_{i=0}^{n} \beta_i X_i = \bm X \bm \beta + \epsilon\]

Where $X$ is the $n \times (p-1)$ design matrix. 

\subsubsection{OLS}
We call the risk under the assumed linear model the \textbf{Residual Sum of Squares}. To estimate the vector of parameters $\bm \beta$ using $\{ (y_i, \bm x_i) \vert i \in [n] \}$ where $x_i = (x_{i1}, x_{i2}, \dots x_{ip})$ represents a row on the design matrix where each $j$ column is a variate (rooms, bathrooms, parking spots) or an attribute (bathrooms squared $x_i^p$) of the $i-$th observation/unit.

\[ RSS(\bm \beta)  := \sum_{i=1}^n(y_i - \beta_0 - \sum_{j=1}^p x_{ij})\]

As a function of the coefficient vector, our estimate for $\bm \hat \beta$ in quadratic matrix form is:

\[ \bm \hat \beta = \argmin_{\bm \beta} (\bm y - \bm X \bm \beta)^T (\bm y - \bm X \bm \beta) \]

The OLS extimate for the coefficient vector is:

\[ \bm \beta_{OLS} = (X^T X)^{-1}X^T\]

We consider the hat matrix $H = X(X^T X)^{-1} X^T$ which:
\begin{itemize}
\item is an \textit{orthogonal projection} matrix since it is indempotent.
\item is indempotent, so  $H^2 = H$
\item the fitted values are the projection of the observed values in the training set y onto the linear subspace spanned by the columnds of \textbf{X}. ie $\hat y = H y$
\item the estimated residuals are the corresponding orthogonal complement $\hat r  = y - Hy = (I-H)y$
\end{itemize}

Properties of the $\tilde{\beta}$


\subsubsection{Assumptions on the residuals}
\begin{itemize}
\item The error terms and thus the $Y_i$ are independent. 
\item The error tersm and thus $Y_i$ are normally distributed (Not uniformly if you want to estimate conditional probabilities for some reason)
\item the error terms and thus $Y_i$ have a constant variance. 
\item the erro terms have a constant mean of 0. Not $Y_i$ necessarily though.n
\end{itemize}

\section{Matrix Differentiation}


\section{Residual Diagnostics}

\subsection{plot of residuals vs fitted values}
Don't shake them up because then you may introduce patterns that don't reallt exist, the fitted values x axis is just for indexing position. 
Examine this by eye to make sure nothing is obviously wrong. Don't look at it too long. This is not enough to determine randomness, or normality of residuals. you will need tests and normal !! plots in addition to this. 

\subsection{ACF}
Auto correlation plot: Will be good for forecasting. Blue bands measure serial correlation. for determining if the residuals are white noise you'd want about 95\% of the spikes to be within the blue lines 











\end{document}
